---
title: "Proof of the slope estimate for the averaging metric"
author: "Xin Jing"
date: "April 30, 2019"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library

```{r}
library(TeachingDemos)
```

## Set seeds

```{r}
char2seed("averaging")
```

## Simulate three ecosystem functions

```{r}
F1 <- rnorm(100, 0, 1)
F2 <- rnorm(100, 0, 1)
F3 <- rnorm(100, 0, 1)
```

## Simulate biodiversity

```{r}
div <- rpois(100, 20)
```

## Calculate the averaging multifunctionality metric

```{r}
Y <- (F1 + F2 + F3)/3
```

## Inspect the variance of the averaging metric 
## and how it comes from the single functions

```{r}
var(Y)
(var(F1) + var(F2) + var(F3) + 2*cov(F1, F2) + 2*cov(F1, F3) + 2*cov(F2, F3))/9
(sd(F1)^2 + sd(F2)^2 + sd(F3)^2 + 2*cov(F1, F2) + 2*cov(F1, F3) + 2*cov(F2, F3))/9
```


## Inspect the covariance and how it comes from the single functions

```{r}
cov(Y, div) 
(cov(F1, div) + cov(F2, div) + cov(F3, div))/3
```

## Calculate the slope between biodiversity and the averaging metric

```{r}
# by the linear model
summary(lm(Y ~ div))
# by the formula
cov(Y, div)/var(div)
# more ways to calculate the slope
cov(Y, div)/cov(div, div)
cor(Y, div)*sd(Y)/sd(div)
```

## Inspect how the slope comes from single functions

```{r}
(cov(F1, div)/cov(div, div) + cov(F2, div)/cov(div, div) + cov(F3, div)/cov(div, div))/3
(cor(F1, div)*sd(F1)/sd(div) + cor(F2, div)*sd(F2)/sd(div) + cor(F3, div)*sd(F3)/sd(div))/3
```

